
#Reproducible Research: Peer Assessment 1

##Loading and preprocessing the data

1. Load the data

```{r}
file <- read.csv("/home/arun/data/assg/RepData_PeerAssessment1/activity.csv", sep = ",", header = TRUE)
```

2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r}
file1 <- aggregate(. ~ date, data=file, FUN=sum, na.rm = TRUE)
```

##Mean of total number of steps taken per day

1.histogram of the total number of steps taken each day

```{r}
library(ggplot2)

qplot(steps, data = file1, fill = date)
```

2.To calculate mean and median  total number of steps taken per day

```{r}
 file_mean <- aggregate(. ~ date, data=file, FUN=mean, na.rm = TRUE)

file_mean

file_median <- aggregate(. ~ date, data=file, FUN=median, na.rm = TRUE)

file_median

```

##Average daily activity pattern

1.  Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}

compl_cases <- subset(file, complete.cases(file) == TRUE)

file_split <- split(compl_cases, compl_cases$interval, drop = TRUE)

interval_Avg1 <- sapply(file_split, function(x){ mean(x$steps)})

plot(interval_Avg1, type="l",  
     main="5 Minute Interval Time Series", 
     ylab="Average Steps", 
     xlab="Interval" )
```

2. 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

```{r}

max <- max(interval_Avg1)

max_pos <- which(interval_Avg1 == max)

max_pos

```

##Imputing missing values

1. Calculation of the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}

#Create subsets with NA`s and without NA`s

compl_cases <- subset(file, complete.cases(file) == TRUE)

compl_cases1 <- subset(file, complete.cases(file) == FALSE)


```

2. strategy for filling in all of the missing values in the dataset.

```{r}

#Replace all NA values by Zero
#This will replace NA`s by '0':

compl_cases1[is.na(compl_cases1)] <- 0

```

3.  new dataset that is equal to the original dataset but with the missing data filled in.

```{r}

new_dataset <- merge(compl_cases, compl_cases1, all = TRUE)

```

4. a)  histogram of the total number of steps taken each day

```{r}

new_dataset1 <- aggregate(. ~ date, data = new_dataset, FUN=sum, na.rm = TRUE)

qplot(steps, data = new_dataset1, fill = date)

```

b) mean and median total number of steps taken per day.

```{r}
new_mean <- aggregate(. ~ date, data=new_dataset1, FUN=mean, na.rm = TRUE)

new_mean

new_median <- aggregate(. ~ date, data=new_dataset1, FUN=median, na.rm = TRUE)

new_median

```

c) difference in values from the estimates from the first part of the assignment and impact of imputing missing data on the estimates of the total daily number of steps

```{r}

summary(new_dataset1)

summary(file1)

par(mfrow = c(1,2))

hist(file1$steps, main="Total Steps per Day", xlab="Steps")

hist(new_dataset1$steps, main="New Total Steps per Day", xlab="Steps")

```

##differences in activity patterns between weekdays and weekends

1. new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day

```{r}
new_dataset$day <- weekdays(as.Date(new_dataset$date))

for (i in 1:nrow(new_dataset)) {                                     
      if (new_dataset[i,]$day %in% c("Saturday","Sunday")) {             
             new_dataset[i,]$day<-"weekend"                                 
         }
       else{
             new_dataset[i,]$day<-"weekday"                                 
        }
}
```

2.  panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}

new_split <- split(new_dataset, new_dataset$interval, drop = TRUE)

Use_plot <- aggregate(new_dataset$steps ~ new_dataset$interval + new_dataset$day, new_dataset, mean)

names(Use_plot) <- c("interval", "day", "steps")

with(Use_plot, plot(steps ~ interval, type = "n", main = "weekday vs weekend"))

with(Use_plot[Use_plot$day == "weekend", ],lines(steps ~ interval, type = "l", col = "blue") )

with(Use_plot[Use_plot$day == "weekday", ],lines(steps ~ interval, type = "l", col = "red") )

legend("topright", lty=c(1,1), col = c("blue", "red"), legend = c("weekend", "weekday"), seg.len=2)
       
```


